On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-11 TTau_Halpha/kurosawa.tex
Corrections made after Tim's comments. The subsection on the fits of T Tau is eliminated from kurosawa.tex.  The subsection was saved in a separate tex file: ttau_model_subsec.tex.  (RK)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@616 25464f21-6099-402c-b30a-640f1d40d38b

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amssymb}

\makeatletter

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{times}

@@

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri stars]{On the formation of H$\alpha$ around classical T Tauri stars}

\author[R. Kurosawa et\,al.]{Ryuichi Kurosawa$^1$\thanks{E-mail: rk@astro.ex.ac.uk}, Tim J. Harries$^1$ and Neil H. Symington$^2$\\
$^1$School of Physics, University of Exeter, Stocker Road, Exeter EX4 4QL. \\
$^2$School of Physics and Astronomy, University of St. Andrews, North Haugh, St. Andrews, Fife, KY16 9SS.}

@@

We present  radiative transfer models of classical T~Tauri stars, and investigate the formation of H$\alpha$
in order to understand their complex circumstellar environment, and to explain
the wide variety of emission line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with magnetospherical accretion alone to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind) presented,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations, 
although the inclination dependence of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al , and discuss the basic physical conditions required to reproduce
the profiles in each classified type. Using the hybrid model, we fit
the observed H$\alpha$ from the optical component of T~Tau triple
system (T~Tau~N), and find its ratio of the mass-loss to mass-accretion
rate to be $\sim0.18$ with the cold ($T_{\mathrm{wind}}\approx8000\,\mathrm{K}$)
bipolar wind model .

@@

T~Tauri stars (TTS) are young ($<\sim 3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass stars, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some
of the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit an excess
amount of continuum flux in the ultraviolet (UV) and infrared (IR).
Their spectral energy distribution and polarisation data suggest the
presence of a circumstellar disc, which plays an important role in
regulating dynamics of gas flows around CTTS.

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidence for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet{takami:2003}
show the direct evidence for the presence of bipolar and monopolar
outflows down to $\sim 1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observation of high-resolution ($R=50\,000$)
two-dimensional spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc, and in both above and below the disc --- suggesting the presence
of the bipolar outflows. On a larger scale, \emph{HST} observation
of HH30 \citep{burrows:1996} shows the jet traced to within $\lesssim30$~AU
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 AU \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their samples (30 CTTS) show blue-shifted
absorption components in at least one of Balmer lines and Ca~K (most
commonly in H$\alpha$).

In the currently favoured model of accretion flows around CTTS, the
accretion discs are disrupted by the magnetosphere of stars which
channels the gas from the disc onto the surface of the stars (e.g.,
\citealt{uchida:1985}; \citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture of the accretion flows is supported
by the evidences that CTTS have relatively strong ($\sim 10^{3}\mathrm{G}$)
magnetic field (e.g., \citealt*{jonhs-krull:1999}; \citealt{guenther:1996})
and by the radiative transfer models which reproduce the observed
profiles for some TTS (\citealt{muzerolle:2001}). The magnetospherical
accretion model naturally explains the blue-ward asymmetric emission
lines (seen in some of CTTS) caused by the partial occultation of
the flow by the disc, and the redshifted absorption component at the
typical free-fall velocities (a few hundred $\mathrm{km\, s^{-1}}$)
seen in some of CTTS. Despite the success of the magnetospherical
accretion model in explaining the line profiles in some CTTS, the
overwhelming observational evidences for the outflow (mentioned above)
in the CTTS profiles suggests that this model is only a part of a
complete picture. Clearly, the modification to include the out-flowing
wind/jet flow is necessary if one requires to predict the mass-accretion
rate and the mass-loss rate of CTTS by modelling their emission profiles
(e.g. H$\alpha$).

Priory to the magnetospherical models, many alternative models had
been considered to explain the observed spectroscopic features mentioned
earlier. For example, (1)~the Alfv\'{e}n wave-driven wind model
(e.g.~\citealt{decampli:1981}; \citealt{hartmann:1982}), (2)~turbulent
boundary layer (between the accretion disc and stellar surface) model
(e.g.~\citealt{bertout:1988}; \citealt{basri:1989}), (3)~chromospheric
model (e.g.~\citealt{calvet:1984}), and (4)~disc wind model (e.g.
\citealt{calvet:1992b}; \citealt{kwan:1995}). Considering the success
of the magnetospherical accretion flow model in some cases and the
observational evidences for the outflows, it is likely that an improved
spectroscopic model requires both inflow and outflow components. The
combination of the magnetospheric accretion flow model with model
1 or 4 would be a reasonable starting point. Although time consuming
to explore a larger parameter space, more realistic density, velocity
and temperature structure from magnetohydrodynamic wind+accretion
models should be considered as inputs for an radiative transfer model
in the future. Recent work by \citet{alencar:2005} showed the observed
H$\alpha$, H$\beta$ and Na~D lines of RW Aur are better produced
by the radiative transfer model which include the collimated disc-wind
arising from near the inner edge of the accretion disc. \citet{malbet:2005}
considered the combination of a spherical wind and the accretion disc
in their radiative transfer model to reproduce Br$\gamma$ from the
early-type Herbig Be star MWC~297.

The magneto-centrifugal wind model, first proposed by \citet{blandford:1982},
has been often used to reproduce the large-scale wind structure of
T Tauri stars, or to model observed optical jets (e.g.~HH~30 jet
by \citealt{burrows:1996}; \citealt{ray:1996}). The launching of
the wind from a Keplerian disc is typically done by treating the equatorial
plane of the disc as a mass-injecting boundary condition (e.g., \citealt{shu:1994};
\citealt{ustyugova:1995}; \citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).
Depending on the location of the open magnetic fields anchored to
the disc, two different types of winds are produced. If the field
is constrained to be near the co-rotation radius of stellar magnetosphere,
an {}``X-wind'' \citep{shu:1994} is produces. If the open field
lines are located in a wider area of the disc, a {}``disc-wind''
similar to \citet{koenigl:2000} is produced \citet{krasnopolsky:2003}.
Recent reviews on the jet/wind-disc connection can be found in \citet{koenigl:2000}
and \citet{pudritz:2005}. Interestingly, \citet{matt:2005} showed
the possibility that the stellar wind along the open magnetic field
originating from the star can cause significant spin-down torque on
the star, provided that mass-loss rate is high enough.

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to perform
the empirical studies of the line formation. More specifically, we
present disc--wind--magnetosphere hybrid models for CTTS, and perform
parametric studies of the H$\alpha$ formation. This study should
provide preliminary physical conditions which lead to the wide variety
of emission line profiles seen in the observation, and would help
to construct more comprehensive circumstellar models (e.g. via MHD
simulations) of T~Tauri stars. Using the model results, we examine
the H$\alpha$ spectroscopic classification used by \citet{reipurth:1996},
and compare models with observations. We will also discuss whether
our model is consistent with some predictions made by the recent MHD
studies i.e.~ $\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model space is divided
into four different regions: (1)~a central continuum source, (2)~magnetosphere
accretion flow, (3)~wind outflow, and (4)~accretion disc. Fig.~\ref{fig:config_dipolar}
depicts the relative location in the model space. In all regions,
the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas freely fall, moving along the magnetic filed onto the
surface of the star. For the purpose of computational simplicity,
the collimated bipolar wind are defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described. In our models, two different
types of the wind configurations are considered: (1)~the bipolar
wind resembling the density structure of the MHD simulation by \citealt{krasnopolsky:2003}
in large scale ($\sim 100$~AU) in section~\ref{sub:wind_jet}, and
(2) that in small scale ($\sim 10$~AU) in section~\ref{sub:Alternative-model-disc-wind}.

@@

\caption{Basic Model Configuration. The system consist of four
components: (1)~the continuum source located at the origin $\left( o
\right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
y-axis are into the paper, (2)~magnetospherical accretion flow,
(3)~bipolar wind outflow, and (4)~accretion disc. The density
distribution is symmetric around the z-axis. The innermost radius of
the magnetosphere (at the equatorial plane) coincides with the the
inner radius of the accretion disc. From the inner most part of the
accretion disc, the gas freely fall, moving along the magnetic filed
onto the surface of the star. For a computational simplicity, the
bipolar wind starts from just outside of the largest radius
($R_{\mathrm{mo}}$) of the magnetosphere (dotted line). }

@@

Unless specified otherwise, we adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, R_{\sun}$, $0.5\, M_{\sun}$, and $4000\,\mathrm{K}$
respectively. The model atmosphere of \citet{kurucz:1979} equivalently
\citet{kurucz:1993}  with $T_{\mathrm{ph}}=4000\,\mathrm{K}$ and
$\log g_{*}=3.5$ (cgs) is used to find the photospheric continuum
at H$\alpha$ wavelength. The parameters are summarised in Table~\ref{tab:std_parameters}.

For a model which includes magnetospherical accretion, an additional
source of the continuum source is considered. As the in-falling gas
becomes closer to the stellar surface, it is decelerates in a strong
shock, and heated to $\sim 10^{6}\,\mathrm{K}$ (with typical parameters).
The X-ray radiation produced in the shock will be absorbed by the
gas located near, and re-emitted in optical and UV radiation (\citealt{koenigl:1991};
\citealt{hartmann:1994}). This will create hot rings near the stellar
surface where the magnetic field intersects with the surface. For
simplicity, the free-falling kinetic energy is assumed to be thermalized
in the radiating layer, and re-emits as blackbody radiation (with
a single temperature). With the parameters of the magnetosphere and
the star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$,
the ratio of this accretion luminosity to the photospheric luminosity
is about $0.5$, and the corresponding temperature of the hot rings is
about $6400\,\mathrm{K}$.  The continuum emission from the hot rings is taken
into account when computing the line profiles.

@@

@@

\caption{The summary of the standard classical T~Tauri star model parameters.}

\label{tab:std_parameters}

@@

We adopt the magnetospherical accretion flow model of \citet{hartmann:1994},
as done so by \citet{muzerolle:2001} and by \citet{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_dipolar}, the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The former value corresponds to the co-rotation radius
of the accretion disc, and the geometry of the magnetic field/stream
lines is kept constant throughout this paper. The geometry of the
magnetosphere is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form: \begin{equation}
r=R_{\mathrm{m}}\sin^{2}\theta\label{eq:dipole}\end{equation}
 \citep[see][]{ghosh:1977} where $r$, and $\theta$ are coordinates
of the field point ($p$) in Fig.~\ref{fig:config_dipolar} in spherical
coordinates, and $R_{\mathrm{m}}$is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$). The range of
$R_{\mathrm{m}}$ is
restricted between $R_{\mathrm{mi}}$ and $R_{\mathrm{mo}}$. Using
the field geometry above and the conservation of energy, the velocity
and the density of the accreting gas along the steam line are found
as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by \citet{hartmann:1994}
is simply adopted here. They computed the temperature assuming a volumetric
heating rate which is proportional to $r^{-3}$, and using the energy
balance of the radiative cooling rate of \citet{hartmann:1982} and
the heating rate \citep{hartmann:1994}. Although the temperature
structure of the accretion steam could significantly affect the line
source function, for the purpose of the exploring the general characteristics
of the H$\alpha$ formation, this simple form is a reasonable assumption.
\citet{martin:1996} presented the self-consistent determination of
the thermal structure of the in-flowing gas along the dipole magnetic
field (equation~\ref{eq:dipole}) by solving the heat equation couple
to rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors of the cooling process are bremsstrahlung radiation
and line emission from Ca~II and Mg~II ions. The results of \citet{martin:1996}
qualitatively agrees with that of \citet{hartmann:1994}.

@@

As mentioned earlier, the wind model presented here is to simulate
the collimated MHD disc-wind model of \citet{krasnopolsky:2003} in
large scale ($\sim 100$~AU) , and the alternative wind model which
resembles the model of \citet{krasnopolsky:2003} in small scale ($\sim
10$~AU)
will be presented later (section~\ref{sub:Alternative-model-disc-wind}).

Similar to the simple wind model of \citet{appenzeller:2005}, the
following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\boldsymbol{v}_{\mathrm{wind}}$, is consist of
radial and azimuthal components which depend on the spherical coordinates
$r$ and $\theta$. The radial component $v_{r}\left(r\right)$ is
assumed to be in the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:

@@

Further, the density of the wind is assumed to be axi-symmetric and
separable in $r$ and $\theta$ for computational simplicity, i.e.,

@@

\caption{The summary of H$\alpha$ equivalent widths from the
magnetospherical accretion flow models show in
Figure~\ref{fig:atlas_acc_models}.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}

\setcounter{tocdepth}{3}
\makeatletter

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

@@

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri
  stars]{On the formation of H$\alpha$ line emission around classical T Tauri stars}

\author[R. Kurosawa et\,al.]{Ryuichi
  Kurosawa$^1$\thanks{E-mail:rk@astro.ex.ac.uk}, Tim J. Harries$^1$
  and Neil H. Symington$^2$\\ $^1$School of Physics, University of Exeter, Stocker Road, Exeter EX4~4QL. \\ $^2$School of Physics and Astronomy, University of St. Andrews, North Haugh, St. Andrews, Fife, KY16 9SS.}

@@

We present  radiative transfer models of classical T~Tauri stars, and
investigate the formation of H$\alpha$ line emission 
in order to understand their complex circumstellar environments, and to explain
the wide variety of line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with magnetospherical accretion alone to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind) presented,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations, 
although the inclination dependence of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al, and discuss the basic physical conditions required to reproduce
the profiles in each classified type.

@@

T~Tauri stars (TTS) are young ($<\sim 3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass objects, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some
of the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit excess
 continuum flux in the ultraviolet (UV) and infrared (IR).
Their spectral energy distribution and polarisation data suggest the
presence of circumstellar discs, which plays an important role in
regulating dynamics of gas flows around CTTS (e.g. \citealt{camenzind:1990}).

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt*{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidence for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt*{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet*{takami:2003}
show direct evidence for the presence of bipolar and monopolar
outflows down to $\sim 1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observations using high-resolution ($R=50\,000$)
two-dimensional spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc both above and below the disc, suggesting the presence
of the bipolar outflows. On a larger scale, \emph{HST} observations
of HH30 \citep{burrows:1996} show the jet traced to within $\lesssim30$~AU
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 AU \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their sample (30 CTTS) show blue-shifted
absorption components in at least one of the Balmer lines and Ca~K (most
commonly in H$\alpha$).

In the currently favoured model of accretion in CTTS, the
accretion disc is disrupted by the magnetosphere of stars which
channels the gas from the disc onto the stellar surface (e.g.,
\citealt{uchida:1985}; \citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture is supported
by the evidence that CTTS have relatively strong ($\sim 10^{3}\mathrm{G}$)
magnetic field (e.g., \citealt{johns-krull:1999};
\citealt{guenther:1996};  \citealt{symington:2005b})
and by radiative transfer models which reproduce the observed
profiles for some TTS (\citealt*{muzerolle:2001}). The magnetospherical
accretion model naturally explains the blue-ward asymmetric emission
lines (seen in some of CTTS) caused by the partial occultation of
the flow by the stellar disc, and the redshifted absorption component at the
typical free-fall velocities (a few hundred $\mathrm{km\, s^{-1}}$)
seen in some of CTTS. Despite the success of the magnetospherical
accretion model in explaining the line profiles in some CTTS, the
overwhelming observational evidence for outflow (mentioned above)
in the CTTS profiles suggests that this model is only a part of the
complete picture. Clearly, a modification to include an out-flowing
wind/jet flow is necessary if one wishes to predict the mass-accretion
rate and the mass-loss rate of CTTS by modelling their emission profiles
(e.g. H$\alpha$).

Prior to the magnetospherical paradigm, many alternative models had
been considered to explain the observed spectroscopic features mentioned
earlier. For example, (1)~the Alfv\'{e}n wave-driven wind model
(e.g.~\citealt{decampli:1981}; \citealt*{hartmann:1982}), (2)~turbulent
boundary layer (between the accretion disc and stellar surface) model
(e.g.~\citealt{bertout:1988}; \citealt{basri:1989}), (3)~chromospheric
model (e.g.~\citealt*{calvet:1984}), and (4)~disc wind model (e.g.
\citealt*{calvet:1992b}; \citealt{kwan:1995}).  Recent work by
\citet{alencar:2005} showed the observed 
H$\alpha$, H$\beta$ and Na~D lines of RW Aur are better be produced
by the radiative transfer model which includes a collimated disc-wind
arising from near the inner edge of the accretion disc. \citet{malbet:2005}
considered a combination of a spherical wind and an accretion disc
in their radiative transfer model to reproduce Br$\gamma$ from the
early-type Herbig Be star MWC~297.

The magneto-centrifugal wind model, first proposed by \citet{blandford:1982},
has been often used to reproduce the large-scale wind structure of
T Tauri stars, or to model observed optical jets (e.g.~HH~30 jet
by \citealt{burrows:1996}; \citealt{ray:1996}). The launching of
the wind from a Keplerian disc is typically done by treating the equatorial
plane of the disc as a mass-injecting boundary condition (e.g., \citealt{shu:1994};
\citealt{ustyugova:1995}; \citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).
Depending on the location of the open magnetic fields anchored to
the disc, two different types of winds are produced. If the field
is constrained to be near the co-rotation radius of stellar magnetosphere,
an {}``X-wind'' \citep{shu:1994} is produces. If the open field
lines are located in a wider area of the disc, a {}``disc-wind''
similar to \citet{koenigl:2000} is produced \citet{krasnopolsky:2003}.
Recent reviews on the jet/wind-disc connection can be found in \citet{koenigl:2000}
and \citet{pudritz:2005}. Interestingly, \citet{matt:2005} showed
the possibility that the stellar wind along the open magnetic field
originated from the star can cause significant spin-down torque on
the star, provided that mass-loss rate is high enough.

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to perform
the empirical studies of the line formation. More specifically, we
present disc--wind--magnetosphere hybrid models for CTTS, and perform
parametric studies of the H$\alpha$ formation. This study should
provide preliminary physical conditions which lead to the wide variety
of emission line profiles seen in the observation, and will help
to construct more comprehensive circumstellar models (e.g. via MHD
simulations) of T~Tauri stars. Using the model results, we examine
the H$\alpha$ spectroscopic classification used by \citet{reipurth:1996},
and compare models with observations. We will also discuss whether
our model is consistent with some predictions made by the recent MHD
studies i.e.~ $\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model space is divided
into four different regions: (1)~a central continuum source, (2)~the magnetospherical
accretion flow, (3)~the wind outflow, and (4)~the accretion
disc. Fig.~\ref{fig:config_dipolar} 
depicts the relative location of the regions in the model space. In
all regions, the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas falls freely, moving along the magnetic field onto the
surface of the star. For simplicity,
the collimated bipolar wind is defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described. In our models, two different
types of the wind configurations are considered: (1)~the bipolar
wind resembling the density structure of the MHD simulation by \citet{krasnopolsky:2003}
in large scale ($\sim 100$~AU) in section~\ref{sub:wind_jet}, and
(2) that in small scale ($\sim 10$~AU) in section~\ref{sub:Alternative-model-disc-wind}.

@@

\caption{The basic model configuration. The system consists of four
components: (1)~the continuum source located at the origin $\left( o
\right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
y-axis are into the paper, (2)~the magnetospherical accretion flow,
(3)~the bipolar wind outflow, and (4)~the accretion disc. The density
distribution is symmetric around the z-axis. The innermost radius of
the magnetosphere (at the equatorial plane) coincides with the the
inner radius of the accretion disc. From the inner most part of the
accretion disc, the gas falls freely, moving along the magnetic field
onto the surface of the star. The bipolar wind starts from just
outside of the largest radius 
($R_{\mathrm{mo}}$) of the magnetosphere (dotted line). }

@@

Unless specified otherwise, we adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, R_{\sun}$, $0.5\, M_{\sun}$, and $4000\,\mathrm{K}$
respectively. The model atmosphere of \citet{kurucz:1993} with
$T_{\mathrm{ph}}=4000\,\mathrm{K}$ and $\log g_{*}=3.5$ (cgs) is used
to find the photospheric continuum at the H$\alpha$ wavelength. The
parameters are summarised in Table~\ref{tab:std_parameters}.

For a model which includes magnetospherical accretion, an additional
continuum source is considered. As the in-falling gas 
becomes close to the stellar surface, it is decelerates in a strong
shock, and is heated to $\sim 10^{6}\,\mathrm{K}$.
The X-ray radiation produced in the shock will be absorbed by the
gas locally, and re-emitted in optical and UV radiation (\citealt{koenigl:1991};
\citealt*{hartmann:1994}). This will create hot rings near the stellar
surface where the magnetic field intersects with the surface. For
simplicity, the free-falling kinetic energy is assumed to be thermalized
in the radiating layer, and re-emits as blackbody radiation (with
a single temperature). With the parameters of the magnetosphere and
the star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$,
the ratio of this accretion luminosity to the photospheric luminosity
is about $0.5$, and the corresponding temperature of the hot rings is
about $6400\,\mathrm{K}$.  The continuum emission from the hot rings is taken
into account when computing the line profiles.

@@

\caption{Summary of the standard classical T~Tauri star model parameters.}
\label{tab:std_parameters}

@@

@@

We adopt the magnetospherical accretion flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_dipolar}, the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The former value corresponds to the co-rotation radius
of the accretion disc, and the geometry of the magnetic field/stream
lines is kept constant throughout this paper. The geometry of the
magnetosphere is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form: \begin{equation}
r=R_{\mathrm{m}}\sin^{2}\theta\label{eq:dipole}\end{equation}
 \citep*[see][]{ghosh:1977} where $r$, and $\theta$ are coordinates
of the field point ($p$) in Fig.~\ref{fig:config_dipolar} in spherical
coordinates, and $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$). The range of
$R_{\mathrm{m}}$ is
restricted between $R_{\mathrm{mi}}$ and $R_{\mathrm{mo}}$. Using
the field geometry above and the conservation of energy, the velocity
and the density of the accreting gas along the steam line are found
as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by \citet{hartmann:1994}
is simply adopted here. They computed the temperature assuming a volumetric
heating rate which is proportional to $r^{-3}$, and using the energy
balance of the radiative cooling rate of \citet{hartmann:1982} and
the heating rate \citep{hartmann:1994}. Although the temperature
structure of the accretion steam could significantly affect the line
source function, for the purpose of the exploring the general characteristics
of the H$\alpha$ formation, this simple form is a reasonable assumption.
\citet{martin:1996} presented the self-consistent determination of
the thermal structure of the in-flowing gas along the dipole magnetic
field (equation~\ref{eq:dipole}) by solving the heat equation couple
to rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors to the cooling process are bremsstrahlung radiation
and line emission from Ca~II and Mg~II ions. The results of \citet{martin:1996}
qualitatively agree with that of \citet{hartmann:1994}.

@@

As mentioned earlier, the wind model presented here is to simulate
the collimated MHD disc-wind model of \citet{krasnopolsky:2003} on a 
large scale ($\sim 100$~AU) , and an alternative wind model which
resembles the model of \citet{krasnopolsky:2003} in small scale ($\sim
10$~AU)
will be presented later (section~\ref{sub:Alternative-model-disc-wind}).

Similar to the simple wind model of \citet{appenzeller:2005}, the
following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\boldsymbol{v}_{\mathrm{wind}}$, consists of
radial and azimuthal components which depend on the spherical coordinates
$r$ and $\theta$. The radial component $v_{r}\left(r\right)$ is
assumed to be in the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:

@@

Further, the density of the wind is assumed to be axi-symmetric and
separable in $r$ and $\theta$,

@@

\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
